Isomorphisms as Groupoids: \\ Homotopy Type Theory with Fractional Types

### 2013-10-13 f.tex
updated f.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Types as Discrete (0-) Groupoids}

@@

In the previous section, every type was modeled by a collection of values
(including a special ``unusable'' value $\bot$) and a collection of trivial
paths. In this section, non-trivial paths will emerge to model ``fractional
values.'' The intuition is that two values that that are connected by a path
should not be counted twice as the path induces an equivalence between these
two values. As illustrated in Sec.~\ref{sec:intuition}, this idea generalizes
to conclude that a value connected to iself using several paths corresponds
to a ``fraction.''

@@

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Quantum Computing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Physics, Information Theory, and Reversible Computing}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Finite Types as Discrete (0-)Groupoids}

@@

In the previous section, every type was modeled by a collection of values
(including a special ``unusable'' value $\bot$) and a collection of trivial
paths. In this section, non-trivial paths will emerge to model ``fractional
values.'' The intuition is that two values that are connected by a path
should not be counted twice as the path induces an equivalence between these
two values: ``they are the same.'' As illustrated in
Sec.~\ref{sec:intuition}, this idea generalizes to conclude that a value
connected to itself using several paths corresponds to a ``fraction.''

@@

There are also deep connections with algebraic types, the seven trees in one,
and the Fiore papers.

@@

Our construction can be interpreted in the context of speculative concurrent
programming. A thread can at any time create, out of nothing, a value of type
$A$ and proceed with it, generating in the process a value of type
$\frac{1}{A}$ that a concurrent process can be manipulate. When an actual
value of type $A$ is produced, the fractional value is used to coordinate a
backtracking mechanism that communicates the actual value of type $A$ to the
original process. This is very similar to how we use logical variables and
unification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Reversible Computing}

We have higher-order reversible computations.

We have constructed a concrete closed monoidal category. We did not use the
Int-construction which requires an existing trace operator. We inherit a lot
of constructions for free: name, coname, trace, etc.

@@

We can interpret $\times$ as parallel computations (i.e., as computations
occurring simultaneously in different regions in space). So a value of type
$A \times B$ consists of two values occupying two different space locations,
e.g., two different memory locations. With that perspective, a value of type
$\frac{1}{A}$ consists of a small dictionary that stores all the values of
type $A$ and can shuffle them around.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Physics and Quantum Computing}

The dual vector space in quantum mechanics and its role in quantum
computing. The fact that $\eta$ and $\epsilon$ are dual is essentially the
teleportation algorithm. We can give it a computational interpretation in
terms of backtracking.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Connections to Information Theory}

We have given classic computational interpretations to negative
information~\cite{negative-information} and negative
entropy~\cite{negative-entropy}.

@@

This is cool and deep.
